Preliminary insights into the use of pottery and culinary practices at Guijiabao site in southwest China

### 2024-04-13 analysis/paper/paper.Rmd
revising based on reviewer comments

The two historical potsherds (ca. 1200-500 BP) from Guijiabao demonstrates mixtures of ruminant adipose and non-ruminant fats, implying a potential increase in the exploitation of C~3~ plants, pigs, or freshwater resources compared to the Neolithic period. In addition to rice, the source of C~3~ plants could also have included wheat, which was introduced into the region during the late Bronze Age (ca. 2,800–2000 cal. BP) [@Huan2022]. For the Gaoshan sample, a slightly enriched δ^13^C may suggest minor contributions of pigs [@Evershed2002], in correspondence with the predominance of domestic pigs at the site [@He2020]. While determining changes in culinary practices over time and regional differences would require a larger sample size, it does contribute to improving our current understanding of direct food practices using organic residues in ceramic vessels. As a preliminary study, our data reveals food contents cooked in pottery from both animal and plant sources, offering valuable insights into culinary practices during the time when agriculture spread into the southwest China.

>>>>

The two historical potsherds (ca. 1200-500 BP) from Guijiabao demonstrates mixtures of ruminant adipose and non-ruminant fats, implying a potential increase in the exploitation of C~3~ plants, pigs, or freshwater resources compared to the Neolithic period. In addition to rice, the source of C~3~ plants could also have included wheat, which was introduced into the region during the late Bronze Age (ca. 2,800–2000 cal. BP) [@Huan2022]. It is surprising that ruminant adipose fat, rather than non-ruminant fat, was detected in the Gaoshan potsherd, given the predominance of domestic pigs in the faunal assemblage at the site [@He2020]. One possibility might be our Gaoshan potsherd is an outlier, potentially not representative of the broader dietary patterns. Another plausible explanation could be that the isotopic values of ruminant adipose fat result from a mixture of non-ruminant adipose fats and ruminant milk fats. This possibility may be supported by the slightly enriched δ^13^C in the Gaoshan sample, suggesting minor contributions from non-ruminant fats, such as those from pigs [@Evershed2002].

While determining changes in culinary practices over time and regional differences would require a larger sample size, it does contribute to improving our current understanding of direct food practices using organic residues in ceramic vessels. As a preliminary study, our data reveals food contents cooked in pottery from both animal and plant sources, offering valuable insights into culinary practices during the time when agriculture spread into the southwest China.
